---
title: "ballots-returned-tracker"

output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(readxl)
library(openxlsx)
library(stringr)

```

```{css, echo=FALSE}

h1.title {
  font-family: roboto;
  color: transparent;
  font-weight: 700;
  text-align: left;
  font-size: 12px;
  padding: 0px;
  height: 0px;
}

.footer {
  #font-family: roboto;
  color: black;
  text-align: left;
  font-size: 12px;
    padding: 5px;
  font-style: italic;
}

h1 {
  #font-family: roboto;
  color: black;
  font-weight: bolder;
  text-align: center;
  font-size: 36px;
  margin-top: 0;
  margin-bottom: 30px;
  
}
h2 {
  #font-family: roboto;
  font-weight: 500;
  color: black;
  text-align: center;
  font-size: 18px;
  margin-top: 0;
  margin-bottom: 2;
}
.bignumber {
  #font-family: roboto;
  color: white;
  font-weight: 900;
  text-align: center;
  font-size: 40px;
  line-height: 65px;
  height: 65px;
  margin-top: 0;
  margin-bottom: 0;
}
body {
  color: black;
  #font-family: roboto;
  font-weight: 400;
  font-size: 18px;
}
h4 {
  font-size: 14px;
  font-style: italic;
  text-align: center;
}
  
   * {
      box-sizing: border-box;
   }
   .card {
      color: white;
      float: left;
      width: calc(20% - 10px);
      padding: 5px;
      border-radius: 10px;
      margin-left: 3px;
      margin-right: 3px;
      margin-top: 3px;
      margin-bottom: 3px;
      height: 100%;
   }
   .card p {
     #font-family: roboto;
     text-align: center;
     font-size: 14px;
     margin-bottom: 0;
   }
   .cardContainer:after {
      content: "";
      display: table;
      clear: both;
   }
   
   @media screen and (max-width: 760px) {
      .bignumber {
         font-size: 32px;
      }
     
     .card p {
         font-size: 13px;
      }


   }
   
   @media screen and (max-width: 650px) {
      .card {
         width: 100%;
      }
      h1.title {
        font-size: 22px;
      }
      
      .bignumber {
         font-size: 35px;
      }
     
     .card p {
         font-size: 14px;
      }
      
      
   }
```

```{r, echo=FALSE, warning=FALSE}


```

```{r, echo=FALSE, warning=FALSE}

counties <- c("Alameda","Alpine","Amador","Butte","Calaveras","Colusa","Contra Costa","Del Norte","El Dorado","Fresno","Glenn","Humboldt","Imperial","Inyo","Kern","Kings","Lake","Lassen","Los Angeles","Madera","Marin","Mariposa","Mendocino","Merced","Modoc","Mono","Monterey","Napa","Nevada","Orange","Placer","Plumas","Riverside","Sacramento","San Benito","San Bernardino","San Diego","San Francisco","San Joaquin","San Luis Obispo","San Mateo","Santa Barbara","Santa Clara","Santa Cruz","Shasta","Sierra","Siskiyou","Solano","Sonoma","Stanislaus","Sutter","Tehama","Trinity","Tulare","Tuolumne","Ventura","Yolo","Yuba")

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

today_UTC <- as.POSIXct(Sys.time()) #get todays date as UTC
today_posix <- format(today_UTC, tz="America/Los_Angeles",usetz=TRUE)
today <- as.Date(substr(as.character(today_posix), 1,10))
yesterday <- as.Date(today-1) #get yesterday's date
filename <- paste(sep="", "vbm-statistics_",yesterday,".xls") #put yesterdays date into filename

#pretty dates
today_display <- format(today, "%A, %b. %d, %Y")

download.file("https://elections.cdn.sos.ca.gov/statewide-elections/2024-general/vbm-statistics.xls", filename) #download file

ballots <- read_excel(filename) #read file

ballots <- ballots %>% 
    row_to_names(row_number = 3)

```

```{r, echo=FALSE, warning=FALSE}

#colnames(ballots)

ballots_clean <- ballots %>% 
  select(`COUNTY`, `Total voters Issued VBM ballots`,`Total Accepted VBM ballots`, `Drop Box`, `Drop Off Location`, `Vote Center Drop Off`, `Mail`, `FAX`, `Other`, `Regular Ballots`, `Provisional Ballots  \n(Non-CVR)`, `Conditional (CVR)`, `Conditional (\"Instant\" CVR )`, `Total Accepted`, `TOTAL BALLOTS CAST`) %>%  #select only the columns we need
  filter(`COUNTY` %in% counties) %>%  #filter out only the rows that have county names
  replace(is.na(.), "0") %>%  #replace everything that's "NULL" with a "0" 
  mutate(`Total voters Issued VBM ballots`= as.numeric(`Total voters Issued VBM ballots`),
         `Total Accepted VBM ballots` = as.numeric(`Total Accepted VBM ballots`), 
         `Drop Box` = as.numeric(`Drop Box`), 
         `Drop Off Location` = as.numeric(`Drop Off Location`), 
         `Vote Center Drop Off` = as.numeric(`Vote Center Drop Off`), 
         `Mail`= as.numeric(`Mail`), 
         `FAX` = as.numeric(`FAX`), 
         `Other` = as.numeric(`Other`),
         `Regular Ballots` = as.numeric(`Regular Ballots`),
         `Provisional Ballots  \n(Non-CVR)` = as.numeric(`Provisional Ballots  \n(Non-CVR)`),
         `Conditional (CVR)` = as.numeric(`Conditional (CVR)`),
         `Conditional (\"Instant\" CVR )` = as.numeric(`Conditional (\"Instant\" CVR )`),
         `Total Accepted` = as.numeric(`Total Accepted`),
         `TOTAL BALLOTS CAST` = as.numeric(`TOTAL BALLOTS CAST`)) %>% #convert them all to numbers
  rename(`VBM Ballots issued`=`Total voters Issued VBM ballots`,
         `VBM Ballots accepted`=`Total Accepted VBM ballots`,
         County = `COUNTY`, 
         `In Person Ballots accepted` =  `Total Accepted`) %>% 
  mutate(`VBM ballots returned` = `Drop Box` + `Drop Off Location` + `Vote Center Drop Off` + `Mail` + `FAX` + `Other`) %>% #get a number for the total VBM ballots returned
  mutate(`In person ballots returned` = `Regular Ballots` + `Provisional Ballots  \n(Non-CVR)` + `Conditional (CVR)` + `Conditional (\"Instant\" CVR )`) %>% #get number for total in person ballots returned 
  mutate(`VBM and In Person Ballots Returned` = `VBM ballots returned` + `In person ballots returned`) %>%  #get total number of ballots returned for both VBM and in person
  mutate(`% VBM returned` = round(((`VBM ballots returned`/`VBM Ballots issued`)*100), digits=2)) %>%  #percent VBM that have been returned
  mutate(`% VBM accepted` = round(((`VBM Ballots accepted`/`VBM Ballots issued`)*100), digits=2)) %>% #percent VBM that have been accepted
#mutate(`% in person accepted` = round(((`In Person Ballots accepted`/`In person ballots returned`)*100), digits=2)) %>%  #percent in person that have been accepted
  mutate(`% total accepted` = round((((`In Person Ballots accepted` + `VBM Ballots accepted`)/(`In person ballots returned` + `VBM Ballots issued`))*100), digits=2))  #percent of total that have been accepted
  #mutate(`% in person accepted` = replace_na(`% in person accepted`, 0))

write.csv(ballots_clean, "ballots_clean.csv", row.names = FALSE)

```

```{r, echo=FALSE, warning=FALSE}

total_ballots_issued = sum(ballots_clean$`VBM Ballots issued`)
total_ballots_returned = sum(ballots_clean$`VBM ballots returned`)
total_ballots_accepted = sum(ballots_clean$`VBM Ballots accepted`)
total_pct_returned = round(((total_ballots_returned/total_ballots_issued)*100), digits=1)
total_pct_accepted = round(((total_ballots_accepted/total_ballots_issued)*100), digits=1)

if (total_ballots_issued > 1000000) {
  total_ballots_issued_display = paste(sep="", as.character(round((total_ballots_issued/1000000), digits=1)), " M")
} else {
  total_ballots_issued_display = paste(sep="", as.character(round((total_ballots_issued/1000), digits=1)), " K")
}


if (total_ballots_returned > 1000000) {
  total_ballots_returned_display = paste(sep="", as.character(round((total_ballots_returned/1000000), digits=1)), " M")
} else {
  total_ballots_returned_display = paste(sep="", as.character(round((total_ballots_returned/1000), digits=1)), " K")
}

if (total_ballots_accepted > 1000000) {
  total_ballots_accepted_display = paste(sep="", as.character(round((total_ballots_accepted/1000000), digits=1)), " M")
} else {
  total_ballots_accepted_display = paste(sep="", as.character(round((total_ballots_accepted/1000), digits=1)), " K")
}


```

<h1>California Ballot Tracker</h1>
<h4>Last updated `r today_display` </h4>

<div class="cardContainer">
<div class="card" style="background-color:#00318b;">
<p>Ballots issued<br>
<span class="bignumber">`r total_ballots_issued_display`<br></span>
</div>
<div class="card" style="background-color:#00318b;">
<p>Ballots returned<br>
<span class="bignumber">`r total_ballots_returned_display`<br></span>
</div>
<div class="card" style="background-color:#00318b;">
<p>Ballots accepted<br>
<span class="bignumber">`r total_ballots_accepted_display`<br></span>
</div>
<div class="card" style="background-color:#00318b;">
<p>% returned<br>
<span class="bignumber">`r total_pct_returned`%<br></span><p>
</div>
<div class="card" style="background-color:#00318b;">
<p>% accepted<br>
<span class="bignumber">`r total_pct_accepted`%<br></span><p>
</div>
</div>

<br>
<br>


<iframe title="Ballots returned and accepted by county" aria-label="Map" id="datawrapper-chart-8PQ2q" src="https://datawrapper.dwcdn.net/Y57wT/1/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="711" data-external="1"></iframe><script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(e){if(void 0!==e.data["datawrapper-height"]){var t=document.querySelectorAll("iframe");for(var a in e.data["datawrapper-height"])for(var r=0;r<t.length;r++){if(t[r].contentWindow===e.source)t[r].style.height=e.data["datawrapper-height"][a]+"px"}}}))}();
</script>

<br>
<br>

<iframe title="Search for your county" aria-label="Dot Plot" id="datawrapper-chart-ycOwi" src="https://datawrapper.dwcdn.net/o6oyq/1/" scrolling="no" frameborder="0" style="width: 0; min-width: 100% !important; border: none;" height="525" data-external="1"></iframe><script type="text/javascript">!function(){"use strict";window.addEventListener("message",(function(e){if(void 0!==e.data["datawrapper-height"]){var t=document.querySelectorAll("iframe");for(var a in e.data["datawrapper-height"])for(var r=0;r<t.length;r++){if(t[r].contentWindow===e.source)t[r].style.height=e.data["datawrapper-height"][a]+"px"}}}))}();
</script>

<br>
<br>
<div class="footer">Built and designed by Lindsey Feingold and Jill Castellano. Source: <a href="https://www.sos.ca.gov/elections/upcoming-elections/general-election-nov-5-2024" target="_blank">California Secretary of State</a>. Data usually lags by about a day.</div>


```{r, echo=FALSE, warning=FALSE}

#trying to get date from header

#x <- XLConnect::loadWorkbook("vbm-statistics.xlsm")
#df <- XLConnect::readWorksheet(x, sheet=1)

#sheet1 <- x$worksheets[[1]]

#headers <- sheet1$headerFooter

```
